---
title: "Shiny App"
format:
  html:
    toc: false
---

```{r}
shiny_app_url <- trimws(Sys.getenv("SHINY_APP_URL", unset = ""))
has_shiny_target <- nzchar(shiny_app_url)
is_local_target <- has_shiny_target && grepl("127\\.0\\.0\\.1|localhost", shiny_app_url)
shiny_target_label <- if (!has_shiny_target) {
  "No target configured"
} else if (is_local_target) {
  "Local target"
} else {
  "Deployment target"
}
shiny_target_note <- if (!has_shiny_target) {
  "No external Shiny URL is configured yet. Use the setup instructions below before publishing the static site."
} else if (is_local_target) {
  "The site currently points to a locally running Shiny process."
} else {
  "The site is currently configured to point to the deployed Shiny application."
}
```

## Open The App

This page is the website entry point for the Shiny application.

`r if (has_shiny_target) paste("Current target:", shiny_app_url) else "Current target: not configured yet."`

::: {.grid-cards}
::: {.card-block}
### Launch Shiny App

`r if (has_shiny_target) sprintf("[Open the Shiny app](%s){target=\"_blank\"}", shiny_app_url) else "Set SHINY_APP_URL before expecting a public launch link here."`

`r shiny_target_note`
:::

::: {.card-block}
### Current Target

- Label: `r shiny_target_label`
- URL: `r if (has_shiny_target) shiny_app_url else "Not configured"`
- Mode: `r if (!has_shiny_target) "Configuration required" else if (is_local_target) "Local development" else "External deployment"`
:::

::: {.card-block}
### Run Command

```r
shiny::runApp("app")
```

Or from the project root:

```powershell
Rscript run_app.R 3838
```
:::
:::

## Embedded Preview

When the target app is available, the frame below gives the website a direct in-page preview.

```{r, echo=FALSE}
if (!has_shiny_target) {
  htmltools::div(
    style = "background:#fff8ef;border:1px solid #eadfce;border-radius:18px;padding:1rem 1.1rem;color:#4f5f68;",
    htmltools::tags$strong("Preview unavailable."),
    htmltools::tags$div("Set SHINY_APP_URL to a deployed app URL, then re-render the site to enable the launch link and embedded preview.")
  )
} else {
  htmltools::div(
    style = "background:#fff;border:1px solid #e3d8c8;border-radius:18px;overflow:hidden;box-shadow:0 12px 28px rgba(27,41,45,0.06);",
    htmltools::div(
      style = "padding:0.9rem 1rem;border-bottom:1px solid #eadfce;background:#fffaf3;font-weight:700;",
      sprintf("Shiny Preview (%s)", shiny_app_url)
    ),
    htmltools::tags$iframe(
      src = shiny_app_url,
      title = "Singapore Tourism Time-Series VA Prototype",
      style = "width:100%;height:900px;border:0;background:#fff;",
      loading = "lazy"
    )
  )
}
```

## Current Module Status

| Module | Current status | Current focus |
|---|---|---|
| Time Series Visual Analysis | lightweight explorer | compare monthly series and metadata before deeper analysis |
| Time Series Clustering | active prototype | Wang branch interaction and visual refinement |
| Forecasting | active prototype | Jin branch forecasting workflow with selectable model comparison and fallback runtime support |

## Data Contract

The current clustering workflow reads from:

- `data/raw/visitor_arrivals_full_dataset.xlsx`
- `data/processed/clustering_country_wide.csv`
- `data/processed/clustering_country_long.csv`
- `data/raw/tourism_update.xlsx` for optional tourism performance context in the forecasting module

## Expected User Flow

1. Open the Shiny app from this page.
2. Enter the `Time Series Clustering` tab.
3. Select a focused set of countries.
4. Choose a year window and normalization mode.
5. Review the outputs in order: `Dashboard`, `Pattern Explorer`, `Focus Market in Context`, `Assignments`.

## Deployment Switch

For local testing, start the app directly with `shiny::runApp("app")` or `Rscript run_app.R 3838`.

When you are ready to switch to `shinyapps.io`, re-render the site with:

```powershell
$env:SHINY_APP_URL='https://YOUR-APP-NAME.shinyapps.io/YOUR-APP/'
quarto render
```

That will update both the launch button and the embedded preview to the deployed app.

## Notes For Local Testing

1. Start the app first if the preview frame is blank.
2. Use the standalone launch link if your browser blocks the embedded frame.
3. Keep the browser window at a normal desktop width when checking layout quality.
