---
title: "Shiny App"
format:
  html:
    toc: false
---

```{r, include=FALSE}
live_shiny_app_url <- "https://zrwang2025mitbsmuedusg.shinyapps.io/singapore-tourism-va/"
shiny_app_url <- trimws(Sys.getenv("SHINY_APP_URL", unset = live_shiny_app_url))
if (!nzchar(shiny_app_url)) {
  shiny_app_url <- live_shiny_app_url
}
has_shiny_target <- nzchar(shiny_app_url)
is_local_target <- has_shiny_target && grepl("127\\.0\\.0\\.1|localhost", shiny_app_url)
shiny_target_label <- if (!has_shiny_target) {
  "No target configured"
} else if (is_local_target) {
  "Local target"
} else {
  "Deployment target"
}
shiny_target_note <- if (!has_shiny_target) {
  "No external Shiny URL is configured yet. Use the setup instructions below before publishing the static site."
} else if (is_local_target) {
  "The site currently points to a locally running Shiny process."
} else {
  "The site is currently configured to point to the deployed Shiny application."
}
```

## Open The App

This page is the website entry point for the Shiny application.

`r if (has_shiny_target) paste("Current target:", shiny_app_url) else "Current target: not configured yet."`

::: {.home-deliverables}
<div class="home-deliverable">
<span>Launch Shiny App</span>
<strong>`r if (has_shiny_target) sprintf("[Open the Shiny app](%s){target=\"_blank\"}", shiny_app_url) else "Set SHINY_APP_URL before expecting a public launch link here."`</strong>
<div>`r shiny_target_note`</div>
</div>
<div class="home-deliverable">
<span>Current Target</span>
<strong>`r shiny_target_label`</strong>
<div>`r if (has_shiny_target) shiny_app_url else "Not configured"`</div>
</div>
<div class="home-deliverable">
<span>User Guide PDF</span>
<strong>[Open the embedded PDF guide](./user-guide.html)</strong>
<div>One final PDF walkthrough for visual analysis, clustering, and forecasting.</div>
</div>
:::

## Embedded Preview

When the target app is available, the frame below gives the website a direct in-page preview.

```{r, echo=FALSE}
if (!has_shiny_target) {
  htmltools::div(
    class = "guide-callout",
    htmltools::tags$strong("Preview unavailable."),
    htmltools::tags$div("Set SHINY_APP_URL to a deployed app URL, then re-render the site to enable the launch link and embedded preview.")
  )
} else {
  htmltools::div(
    style = "background:#fff;border:1px solid #e3d8c8;border-radius:20px;overflow:hidden;box-shadow:0 16px 30px rgba(27,41,45,0.08);",
    htmltools::div(
      style = "padding:0.9rem 1rem;border-bottom:1px solid #eadfce;background:#fffaf3;font-weight:700;",
      sprintf("Shiny Preview (%s)", shiny_app_url)
    ),
    htmltools::tags$iframe(
      src = shiny_app_url,
      title = "Singapore Tourism Time-Series VA Prototype",
      style = "width:100%;height:900px;border:0;background:#fff;",
      loading = "lazy"
    )
  )
}
```

## Current Module Status

| Module | Current status | Current focus |
|---|---|---|
| Time Series Visual Analysis | integrated EDA dashboard | map source-market concentration and compare pre-COVID versus recovery leaders |
| Time Series Clustering | guided workspace | compare recovery trajectories, recovery position, focus-market placement, and export assignments |
| Forecasting | forecasting studio | compare model families, inspect diagnostics, and relate arrivals forecasts to tourism-performance context |

## Expected User Flow

Start in `Time Series Visual Analysis` to understand the market structure. Move to `Time Series Clustering` to compare country recovery trajectories and isolate one priority market. Finish in `Forecasting` to compare model performance and interpret the forward projection with context indicators. The detailed workflow is documented on the [User Guide PDF](./user-guide.html) page.

## Data Contract

The current app reads from:

- `data/raw/visitor_arrivals_full_dataset.xlsx`
- `data/processed/arrivals_country_wide.csv`
- `data/processed/arrivals_country_long.csv`
- `data/processed/clustering_country_wide.csv`
- `data/processed/clustering_country_long.csv`
- `data/processed/clustering_series_metadata.csv`
- `data/raw/tourism_update.xlsx` for tourism-performance context

## Local Run

For local testing, start the app directly with `shiny::runApp("app")` or `Rscript run_app.R 3838`.

If the deployment target ever changes, re-render the site with:

```powershell
$env:SHINY_APP_URL='https://YOUR-APP-NAME.shinyapps.io/YOUR-APP/'
quarto render
```

That updates both the launch button and the embedded preview to the deployed app.
