---
title: "Singapore Tourism Recovery Visual Analytics CDA"
format: html
author: "Xi Zixun"
---

The exploratory data analysis (EDA) highlighted several structural patterns in Singapore’s tourism data, including the pandemic shock to visitor arrivals, changes in source-market importance, and possible seasonal differences. Building on these observations, confirmatory data analysis (CDA) is used to statistically test whether these patterns are supported by the data.

## Preparation

In this analysis, three tourism periods are defined:

- **Pre-COVID**: 2017 to 2019
- **COVID Shock**: 2020 to 2021
- **Recovery**: 2022 to 2025

### Package Loading

- `tidyverse` for data wrangling and static visualisation
- `lubridate` for date handling
- `plotly` for interactive charts
- `patchwork` for combining ggplot figures
- `GGally` and `corrplot` for exploratory relationship checks
- `readxl` for reading the Excel workbook
- `psych` for descriptive statistics

```{r}
pacman::p_load(tidyverse, lubridate, plotly, patchwork, GGally, corrplot, readxl, psych)
```

## Data Loading

```{r}
path <- "../data/raw/tourism_update.xlsx"

tour <- read_excel(
  path,
  sheet = "My Series",
  skip = 29,
  col_names = FALSE,
  .name_repair = "minimal"
)

names(tour) <- c(
  "Date",
  "Visitor Arrivals",
  "Tourism Receipts: YTD: YoY",
  "Tourist Expenditure Per Capita",
  "Visitor Arrivals: China",
  "Average Length of Stay",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Number of Hotels",
  "Visitor Days",
  "Hotel Room Occupancy Rate",
  "Visitor Arrivals: West Asia",
  "Visitor Arrivals: Taiwan",
  "Total Room Revenue",
  "Number of Hotels.1",
  "Average Length of Stay.1",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Hotel Revenue: per Available Room: Luxury",
  "Hotel Revenue: per Available Room: Mid-Tier",
  "Visitor Arrivals: ASEAN",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "No of Hotel Room Stock",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Africa",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: North Asia",
  "Visitor Arrivals: 8-10 Days",
  "Visitor Arrivals: 11-14 Days",
  "Visitor Arrivals: 15 Days & Over",
  "Visitor Arrivals: Americas",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour <- tour %>%
  mutate(Date = as.Date(Date))

glimpse(tour)
```

## Data Cleaning

```{r}
data_clean <- tour %>%
  filter(year(Date) >= 2017 & year(Date) <= 2025) %>%
  mutate(
    Year = year(Date),
    Month = month(Date),
    Period = case_when(
      Year >= 2017 & Year <= 2019 ~ "Pre-COVID",
      Year >= 2020 & Year <= 2021 ~ "COVID Shock",
      Year >= 2022 & Year <= 2025 ~ "Recovery"
    ),
    Period = factor(Period, levels = c("Pre-COVID", "COVID Shock", "Recovery")),
    China_Share = (`Visitor Arrivals: China` / `Visitor Arrivals`) * 100
  )

data_clean %>% count(Year, Month)
summary(data_clean$`Visitor Arrivals`)
```

## Descriptive Analysis

```{r}
describe(
  data_clean %>%
    select(`Visitor Arrivals`, `Visitor Arrivals: China`, China_Share,
           `Hotel Room Occupancy Rate`, `Average Length of Stay`, `Visitor Days`)
)
```

## CDA (Confirmatory Data Analysis)

### CDA 1 - Statistical Impact of COVID-19 on Visitor Arrivals

Did the COVID-19 pandemic significantly affect visitor arrivals to Singapore?

**Hypothesis**

H₀ (Null Hypothesis)  
There is no significant difference in visitor arrivals across the tourism periods.

H₁ (Alternative Hypothesis)  
Visitor arrivals differ significantly across the tourism periods.

```{r}
plot_ly(
  data = data_clean,
  x = ~Period,
  y = ~`Visitor Arrivals`,
  color = ~Period,
  type = "box",
  boxpoints = "outliers"
) %>%
  layout(
    title = "Visitor Arrivals Across Tourism Periods",
    xaxis = list(title = "Period"),
    yaxis = list(title = "Visitor Arrivals"),
    showlegend = FALSE
  )
```

```{r}
anova_model <- aov(`Visitor Arrivals` ~ Period, data = data_clean)
summary(anova_model)
```

::: callout-note
The boxplot compares the distribution of overall visitor arrivals across the three tourism periods. A marked downward shift is expected during the COVID Shock period relative to the Pre-COVID baseline, followed by a partial rebound in the Recovery period. If the ANOVA result is statistically significant, this supports the conclusion that visitor arrivals differed meaningfully across the three phases of tourism activity.
:::

### CDA 2 - Influence of China Share on Total Visitor Arrivals

Does China’s share of Singapore’s inbound tourism market significantly influence total visitor arrivals?

**Hypothesis**

H₀ (Null Hypothesis)  
China’s visitor share has no significant relationship with total visitor arrivals.

H₁ (Alternative Hypothesis)  
China’s visitor share significantly influences total visitor arrivals.

```{r}
ggplot(data_clean, aes(x = China_Share, y = `Visitor Arrivals`)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Relationship Between China Market Share and Total Visitor Arrivals",
    x = "China Share of Total Visitor Arrivals (%)",
    y = "Visitor Arrivals"
  ) +
  theme_minimal()
```

```{r}
model_cda2 <- lm(`Visitor Arrivals` ~ China_Share, data = data_clean)
summary(model_cda2)
```

::: callout-tip
Interpret the regression using both the p-value and the $R^2$ value. A statistically insignificant coefficient would suggest that changes in China’s share alone do not explain much of the variation in Singapore’s total visitor arrivals.
:::

### CDA 3 - Seasonal Variation in Visitor Arrivals

Do visitor arrivals vary significantly across months?

**Hypothesis**

H₀ (Null Hypothesis)  
Visitor arrivals do not differ significantly across months.

H₁ (Alternative Hypothesis)  
Visitor arrivals vary significantly across months, indicating seasonal tourism patterns.

```{r}
data_no_covid <- data_clean %>%
  filter(!Year %in% c(2020, 2021)) %>%
  mutate(Month_Label = month(Date, label = TRUE, abbr = TRUE))
```

```{r}
p <- ggplot(
  data_no_covid,
  aes(
    x = Month_Label,
    y = `Visitor Arrivals`,
    text = paste(
      "Month:", Month_Label,
      "<br>Visitor Arrivals:", format(`Visitor Arrivals`, big.mark = ",")
    )
  )
) +
  geom_boxplot(fill = "skyblue") +
  labs(
    title = "Seasonal Variation in Visitor Arrivals",
    x = "Month",
    y = "Visitor Arrivals"
  ) +
  theme_minimal()

ggplotly(p, tooltip = "text")
```

```{r}
anova_month <- aov(`Visitor Arrivals` ~ Month_Label, data = data_no_covid)
summary(anova_month)
```

::: callout-tip
The monthly distributions are based on observations from 2017–2025, with the COVID shock years (2020–2021) excluded to reduce distortion in seasonal patterns.
:::

A statistically significant ANOVA result would indicate that average visitor arrivals differ across months, suggesting seasonality in tourism demand. If the p-value is not significant, this would imply limited evidence of strong seasonal differences in the dataset after excluding the COVID shock years.
