---
title: "Singapore Tourism Recovery Visual Analytics EDA"
format: html
author: "Xi Zixun"
---

## Background

Singapore’s inbound tourism landscape is characterised by a concentrated set of major source markets and a geographical distribution that evolves over time. Analysis of visitor arrivals from 2017 to 2025 indicates that Singapore’s tourism demand is anchored primarily in the Asia-Pacific region, while the relative importance of individual source countries shifts across different years. Such variation is visible not only in country rankings and market shares, but also in the wider spatial pattern of visitor origins. These dynamics make it insufficient to rely on static summary figures alone. A more effective approach is to integrate geospatial and comparative visual analysis in order to examine both the global distribution of visitor origins and the changing prominence of leading markets. Accordingly, this project proposes an interactive visual analytics prototype to support a clearer understanding of where Singapore’s visitors come from and how the structure of its source markets changes over time.

## ndefinedPreparation

### Package Installation & Loading

-   `tidyverse`- A family of modern R packages specially designed to support data science, analysis and communication task including creating static statistical graphs.
-   `plotly` - R library for plotting interactive statistical graphs.
-   `patchwork` - combining multiple ggplot2 graphs into one figure.
-   `GGally` - 'GGally' extends 'ggplot2' by adding several functions to reduce the complexity of combining geometric objects with transformed data
-   `corrplot` - a visual exploratory tool on correlation matrix that supports automatic variable reordering to help detect hidden patterns among variables.
-   `readx1` - Graphical scales map data to aesthetics, and provide methods for automatically determining breaks and labels for axes and legends.
-   `lubridate` - provides intuitive functions for parsing, manipulating, and performing arithmetic on dates, times, and time zones

```{r}
pacman::p_load(tidyverse, lubridate, plotly, patchwork, GGally, corrplot, readxl, cellranger, knitr, scales)
```

## Data Exploration & Cleaning

```{r}
path <- "../data/raw/tourism_update.xlsx"

tour <- read_excel(
  path,
  sheet = "My Series",
  skip = 29,
  col_names = FALSE,
  .name_repair = "minimal"
)

names(tour) <- c(
  "Date",
  "Visitor Arrivals",
  "Tourism Receipts: YTD: YoY",
  "Tourist Expenditure Per Capita",
  "Visitor Arrivals: China",
  "Average Length of Stay",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Number of Hotels",
  "Visitor Days",
  "Hotel Room Occupancy Rate",
  "Visitor Arrivals: West Asia",
  "Visitor Arrivals: Taiwan",
  "Total Room Revenue",
  "Number of Hotels.1",
  "Average Length of Stay.1",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Hotel Revenue: per Available Room: Luxury",
  "Hotel Revenue: per Available Room: Mid-Tier",
  "Visitor Arrivals: ASEAN",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "No of Hotel Room Stock",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Africa",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: North Asia",
  "Visitor Arrivals: 8-10 Days",
  "Visitor Arrivals: 11-14 Days",
  "Visitor Arrivals: 15 Days & Over",
  "Visitor Arrivals: Americas",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour <- tour %>%
  mutate(Date = as.Date(Date))

tour_clean <- tour %>%
  filter(year(Date) >= 2017, year(Date) != 2026)
```

```{r}
names(tour)
head(tour$Date)
class(tour$Date)
unique(year(tour_clean$Date))
```

::: callout-tip
Some preparation of data to be noticed:

-   The year 2015, 2016 and 2026 were removed because they do not consist of full year data from **Jan to Dec**
:::

```{r}
country_cols <- c(
  "Visitor Arrivals: China",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Visitor Arrivals: Taiwan",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour_long <- tour_clean %>%
  select(Date, all_of(country_cols)) %>%
  pivot_longer(
    cols = -Date,
    names_to = "Country",
    values_to = "Arrivals"
  ) %>%
  mutate(
    Country = str_remove(Country, "^Visitor Arrivals: ")
  )
```

::: callout-note
The dataset was reshaped into long format so that each row represents one country’s visitor arrivals for one specific month. The resulting structure contains three main variables: `Date`, `Country`, and `Arrivals`. This format is more suitable for time-series visualization and comparative analysis in R, as it allows countries to be treated as categories within a single variable rather than as separate columns. As a result, it supports cleaner code, easier grouping and filtering, and more effective use of ggplot2 for visualizing changes in visitor arrivals over time.
:::

Before we continue with analysis results, some Date periods should be noticed: **AY2017 Start** - **AY2019 End** is consider as pre_covid period, **AY2020 Start** - **AY2021 End** is Consider as covid_shock, **AY2022 Star**t - **AY2025 End** is considered as recovery period

## EDA 1 - Singapore Visitors

```{r}
country_cols <- c(
  "Visitor Arrivals: China",
  "Visitor Arrivals: Malaysia",
  "Visitor Arrivals: India",
  "Visitor Arrivals: Indonesia",
  "Visitor Arrivals: Australia",
  "Visitor Arrivals: Taiwan",
  "Visitor Arrivals: Hong Kong SAR (China)",
  "Visitor Arrivals: Italy",
  "Visitor Arrivals: Russian Federal (CIS)",
  "Visitor Arrivals: France",
  "Visitor Arrivals: Philippines",
  "Visitor Arrivals: Spain",
  "Visitor Arrivals: Thailand",
  "Visitor Arrivals: Ireland",
  "Visitor Arrivals: United Arab Emirates",
  "Visitor Arrivals: United Kingdom",
  "Visitor Arrivals: Bangladesh",
  "Visitor Arrivals: Iran",
  "Visitor Arrivals: New Zealand",
  "Visitor Arrivals: Israel",
  "Visitor Arrivals: Germany",
  "Visitor Arrivals: Scandinavia: Sweden",
  "Visitor Arrivals: Switzerland",
  "Visitor Arrivals: USA",
  "Visitor Arrivals: Canada",
  "Visitor Arrivals: Mauritius",
  "Visitor Arrivals: Kuwait",
  "Visitor Arrivals: Egypt",
  "Visitor Arrivals: Brunei",
  "Visitor Arrivals: Finland",
  "Visitor Arrivals: Japan",
  "Visitor Arrivals: South Korea",
  "Visitor Arrivals: Myanmar",
  "Visitor Arrivals: Netherlands",
  "Visitor Arrivals: Scandinavia: Norway",
  "Visitor Arrivals: Saudi Arabia",
  "Visitor Arrivals: Sri Lanka",
  "Visitor Arrivals: Vietnam",
  "Visitor Arrivals: Pakistan",
  "Visitor Arrivals: Republic of South Africa"
)

tour_long <- tour_clean %>%
  select(Date, all_of(country_cols)) %>%
  pivot_longer(
    cols = -Date,
    names_to = "Country",
    values_to = "Arrivals"
  ) %>%
  mutate(
    Country = str_remove(Country, "^Visitor Arrivals: "),
    Year = year(Date)
  )
```

```{r}
geo_year <- tour_long %>%
  group_by(Year, Country) %>%
  summarise(
    Arrivals = sum(Arrivals, na.rm = TRUE),
    .groups = "drop"
  )

geo_year <- geo_year %>%
  mutate(
    Country_map = recode(
      Country,
      "USA" = "United States",
      "Russian Federal (CIS)" = "Russia",
      "Hong Kong SAR (China)" = "Hong Kong",
      "Scandinavia: Sweden" = "Sweden",
      "Scandinavia: Norway" = "Norway",
      "Republic of South Africa" = "South Africa"
    )
  )

```

```{r}
plot_ly(
  data = geo_year,
  type = "choropleth",
  locations = ~Country_map,
  locationmode = "country names",
  z = ~Arrivals,
  frame = ~Year,
  text = ~paste(
    "Country:", Country,
    "<br>Year:", Year,
    "<br>Arrivals:", format(Arrivals, big.mark = ",")
  ),
  hoverinfo = "text",
  colorscale = "Reds"
) %>%
  layout(
    title = "Visitor Arrivals to Singapore by Source Country (2017-2025)",
    geo = list(showframe = FALSE, showcoastlines = TRUE)
  )

```

The geospatial results indicate that Singapore’s visitor arrivals are heavily concentrated in Asia, with nearby regional markets contributing the largest share of tourism inflows. This highlights the continued importance of geographical proximity, regional connectivity, and established travel links in shaping visitor demand. The map also shows that arrivals are not evenly distributed across the world, but are instead clustered around a limited number of key source markets.

Over time, the spatial pattern changes noticeably, especially around the pandemic period. Visitor distribution weakens during the years of disruption and gradually strengthens again in the later years, although the pace of recovery appears uneven across countries. This suggests that while Singapore’s tourism sector has recovered, the composition of its visitor markets may not have returned uniformly to its earlier pattern. Overall, the findings show that Singapore’s tourism demand remains regionally concentrated, but also sensitive to major external disruptions.

## EDA 2 - Top Visitor Countries Comparison Pre and Post Covid

```{r}
p1_data <- tour_long %>%
  filter(year(Date) >= 2017 & year(Date) <= 2019)

p2_data <- tour_long %>%
  filter(year(Date) >= 2022 & year(Date) <= 2025)
```

::: panel-tabset
## Table - Pre_Covid

```{r}
p1_rank <- p1_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    Share_Percent = (Avg_Monthly_Arrivals / sum(Avg_Monthly_Arrivals)) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

p1_rank
```

## Chart - Pre_Covid

```{r}
p1_rank <- p1_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE), 0),
    .groups = "drop"
  ) %>%
  mutate(
    Share_Percent = (Avg_Monthly_Arrivals / sum(Avg_Monthly_Arrivals)) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

plot_ly(
  data = p1_rank,
  x = ~Avg_Monthly_Arrivals,
  y = ~reorder(Country, Avg_Monthly_Arrivals),
  type = "bar",
  orientation = "h",
  text = ~paste0(round(Share_Percent, 1), "%"),
  textposition = "outside",
  hovertemplate = "Average Monthly Arrivals: %{x:,}<extra></extra>"
) %>%
  layout(
    title = "P1 (2017-2019): Top 5 Visitor Countries",
    xaxis = list(title = "Average Monthly Arrivals"),
    yaxis = list(title = "Country")
  )
```
:::

In the pre-COVID period (2017–2019), Singapore’s top visitor source countries were led by China and Indonesia, which recorded average monthly arrivals of 285,357 and 252,402 respectively. Their corresponding shares, 19.35% and 17.12%, show that these two countries formed the dominant core of Singapore’s inbound tourism market. India, Malaysia, and Australia followed at a much lower level, with average monthly arrivals of 114,787, 101,198, and 92,571 respectively.

These findings suggest that Singapore’s visitor market before COVID-19 was concentrated in a relatively small number of major source countries, with particularly strong dependence on nearby Asian markets. The ranking pattern reflects the importance of regional accessibility and established travel demand, and provides a useful baseline for evaluating whether the relative importance of these markets changed in the later recovery period.

::: panel-tabset
## Table - Post_Covid

```{r}
p2_rank <- p2_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    Share_Percent = (Avg_Monthly_Arrivals / sum(Avg_Monthly_Arrivals)) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

p2_rank
```

## Chart - Post_Covid

```{r}
p2_rank <- p2_data %>%
  group_by(Country) %>%
  summarise(
    Avg_Monthly_Arrivals = round(mean(Arrivals, na.rm = TRUE), 0),
    .groups = "drop"
  ) %>%
  mutate(
    Share_Percent = (Avg_Monthly_Arrivals / sum(Avg_Monthly_Arrivals)) * 100
  ) %>%
  arrange(desc(Avg_Monthly_Arrivals)) %>%
  mutate(Rank = row_number()) %>%
  slice_head(n = 5)

plot_ly(
  data = p2_rank,
  x = ~Avg_Monthly_Arrivals,
  y = ~reorder(Country, Avg_Monthly_Arrivals),
  type = "bar",
  orientation = "h",
  text = ~paste0(round(Share_Percent, 1), "%"),
  textposition = "outside",
  hovertemplate = "Average Monthly Arrivals: %{x:,}<extra></extra>"
) %>%
  layout(
    title = "P2 (2022-2025): Top 5 Visitor Countries",
    xaxis = list(title = "Average Monthly Arrivals"),
    yaxis = list(title = "Country")
  )
```
:::

The P2 results, covering the post-COVID recovery period from 2022 to 2025, show that Singapore’s visitor market recovered with a structure that was broadly similar to the pre-pandemic period, but with some notable shifts in relative ranking and scale. Indonesia ranked first during this period with an average monthly arrival figure of 173,636, accounting for 16.43% of the total arrivals among the countries included in the analysis. China followed in second place with 159,896 average monthly arrivals and a share of 15.05%. This indicates that although both markets remained central to Singapore’s inbound tourism recovery, Indonesia overtook China in the post-pandemic period.

India remained in third place with an average monthly arrival level of 86,675 and a share of 8.16%, while Malaysia and Australia followed closely with 86,258 (8.12%) and 85,258 (8.02%) respectively. Compared with the pre-COVID period, the ranking composition of the top five markets remained unchanged, which suggests that Singapore’s core visitor base showed a relatively high degree of resilience. However, the reduced arrival levels across all five countries indicate that recovery was incomplete in absolute terms, even though the main source markets continued to dominate.

A clearer view of recovery emerges when P2 is compared directly with the pre-COVID baseline. China’s average monthly arrivals fell from 285,357 in P1 to 159,896 in P2, while Indonesia declined from 252,402 to 173,636. Similar declines are also visible for India, Malaysia, and Australia. This suggests that by 2022–2025, Singapore’s tourism sector had regained its core market structure, but not yet its previous scale. In other words, recovery appears to have been stronger in terms of market composition than in terms of visitor volume.

At the same time, the ranking reversal between China and Indonesia is an important signal. It suggests that not all major source markets recovered at the same speed. Indonesia’s stronger post-pandemic position may reflect a faster rebound in short-haul regional travel, while China’s fall to second place suggests a relatively slower recovery compared with its pre-COVID dominance. The fact that India, Malaysia, and Australia remained in the top five also points to continued dependence on established Asia-Pacific markets, reinforcing the regional nature of Singapore’s tourism recovery.
