---
title: "Module Prototype: Forecasting"
execute:
  warning: false
  message: false
---

```{r}
source("../app/R/data_utils.R")
library(dplyr)
library(ggplot2)
library(DT)
library(lubridate)
options(scipen = 999)
```

## Objective

This model first understands the monthly time series pattern, then builds and compares predictive models on the holdout window, and finally refits the best model to generate forward-looking predictions. The general dataset is organized around monthly visitor arrivals by country, while hotel occupancy rates, average length of stay, and room revenue are used as auxiliary indicators to interpret whether the demand recovery is reflected in tourism performance.

## Data Contract

Shared arrivals backbone:

- `data/raw/visitor_arrivals_full_dataset.xlsx`

Optional supporting tourism context:

- `data/raw/tourism_update.xlsx`

This prediction prototype narrows the scope of analysis to the monthly visitor arrival sequences by country. Meanwhile, hotel occupancy rates, length of stay, hotel numbers and room revenue will be available as optional supplementary background information.

Examples of eligible target series are:

1. `Visitor Arrivals: China`
2. `Visitor Arrivals: Malaysia`
3. `Visitor Arrivals: India`
4. `Visitor Arrivals: Indonesia`
5. `Visitor Arrivals: Australia`
6. `Visitor Arrivals: Japan`

```{r}
tourism_data <- load_tourism_data()
series_catalog <- list_country_arrival_series(tourism_data$long_monthly)
head(series_catalog, 10)
```

## Runtime Strategy

```{r}
stack_status <- forecast_stack_status()

tibble(
  fallback_ready = stack_status$fallback_ready,
  modeltime_ready = stack_status$modeltime_ready,
  preferred_engine = stack_status$preferred_engine,
  missing_modeltime_packages = ifelse(
    length(stack_status$missing_modeltime_packages) == 0,
    "None",
    paste(stack_status$missing_modeltime_packages, collapse = ", ")
  )
)
```

## Analytical Framing

The forecasting module uses a two-layer logic:

1. **Core forecasting series**: country-level monthly visitor arrivals.
2. **Supporting performance indicators**: hotel room occupancy rate, average length of stay, and total room revenue.

This means the forecasts answer the question *"how may demand from each source market evolve"* while the supporting indicators help answer *"does that demand recovery translate into broader tourism performance"*

## Forecasting Workflow

This model will be divided into seven steps：

1. Import and inspect the selected monthly country-arrivals series.
2. Compare that country series with hotel and stay indicators.
3. Visualise the time path and seasonal structure.
4. Split the series into training and testing sets.
5. Fit a baseline and multiple forecasting models.
6. Compare testing-set accuracy.
7. Refit the best model and forecast forward.

## Example Target Series

`Visitor Arrivals: China` is used here because it is one of the clearest country-level recovery indicators in the dataset and it shows strong shock, rebound, and seasonal dynamics.

```{r}
example_label <- "Visitor Arrivals: China"

example_series <- prepare_forecast_series(
  tourism_data$long_monthly,
  example_label
)

forecast_results <- run_forecast_workflow(
  series_df = example_series,
  horizon = 12,
  engine = "auto"
)

context_series <- prepare_country_context_panel(
  tourism_data$long_monthly,
  country_label = example_label
)
```

For this prototype run, the selected execution engine is **`r forecast_results$engine_label`**.

## Step 1: Position the Country Series Within Tourism Performance

Before forecasting, the selected country-arrivals series should be interpreted together with hotel and stay indicators. The values are normalized so the focus stays on **shared turning points** rather than raw units.

```{r}
ggplot(context_series, aes(x = date, y = normalized_value, color = label)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Country Arrivals Compared with Supporting Tourism Indicators",
    subtitle = "Normalized z-scores show whether demand recovery aligns with occupancy, stay length, and room revenue",
    x = NULL,
    y = "Normalized z-score",
    color = NULL
  ) +
  theme_minimal(base_size = 13)
```

## Step 2: Visualise the Raw Time Series

The chart below shows the full monthly path of the selected country-arrivals series.

```{r}
ggplot(example_series, aes(x = date, y = value)) +
  geom_line(linewidth = 1, color = "#0f6b6f") +
  geom_point(size = 1.8, color = "#d86f45") +
  labs(
    title = example_label,
    subtitle = "Monthly country-level visitor arrivals used for forecasting",
    x = NULL,
    y = "Visitor arrivals (person)"
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_minimal(base_size = 13)
```

## Step 3: Check Seasonality and Decomposition

The next step is to confirm whether the series contains strong seasonal cycles and how the trend changed around the shock-and-recovery period. 

### Seasonal Pattern by Month

```{r}
example_series |>
  mutate(
    month_lab = month(date, label = TRUE, abbr = TRUE),
    year_num = year(date)
  ) |>
  ggplot(aes(x = month_lab, y = value, group = year_num, color = factor(year_num))) +
  geom_line(linewidth = 0.8, alpha = 0.65) +
  geom_point(size = 1.3, alpha = 0.8) +
  labs(
    title = "Seasonal Comparison by Month",
    subtitle = "Each coloured line represents one year",
    x = NULL,
    y = "Visitor arrivals (person)",
    color = "Year"
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none")
```

### STL-style Decomposition

```{r}
example_ts <- ts(
  example_series$value,
  start = c(year(min(example_series$date)), month(min(example_series$date))),
  frequency = 12
)

decomp_tbl <- stats::stl(example_ts, s.window = "periodic")

forecast::autoplot(decomp_tbl) +
  labs(
    title = "Trend / Seasonal / Remainder Decomposition",
    subtitle = "Used to explain the structural change before forecasting"
  )
```

## Step 4: Create the Training and Testing Split

The final 12 months are reserved as a holdout set. This keeps the evaluation time-aware and avoids random sampling.

```{r}
split_summary <- tibble(
  segment = c("Training", "Testing"),
  start = c(min(forecast_results$training$date), min(forecast_results$testing$date)),
  end = c(max(forecast_results$training$date), max(forecast_results$testing$date)),
  n_obs = c(nrow(forecast_results$training), nrow(forecast_results$testing))
)

split_summary
```

## Step 5: Fit Baseline and Forecasting Models

This prototype includes:

1. `Seasonal Naive` as the baseline benchmark.
2. `ETS (Modeltime)` using exponential smoothing.
3. `ARIMA` using `auto_arima` through the modeltime workflow.


```{r}
forecast_results$models_tbl
```

## Step 6: Testing-Set Forecast and Accuracy Comparison

### Accuracy Table

```{r}
DT::datatable(
  forecast_results$accuracy_tbl,
  rownames = FALSE,
  options = list(dom = "t", pageLength = 6, scrollX = TRUE)
)
```

### Testing Window Forecast Plot

```{r}
plot_forecast_results(forecast_results, type = "holdout") +
  labs(
    title = "Testing-Set Forecast Comparison",
    subtitle = paste(
      "Seasonal naive is the benchmark; ETS and ARIMA are model-based alternatives | Engine:",
      forecast_results$engine_label
    )
  )
```

## Step 7: Refit the Best Model and Forecast Forward

After accuracy comparison, the best-performing model is refit on the full series and projected forward for the next 12 months. 

```{r}
plot_forecast_results(forecast_results, type = "future") +
  labs(
    title = "Forward Forecast After Refit",
    subtitle = paste("Best model refit on the full monthly series | Engine:", forecast_results$engine_label)
  )
```

## Interpretation Notes

1. It starts with time-series exploration rather than jumping straight to prediction.
2. It treats the data as an ordered monthly sequence.
3. It uses a proper holdout split based on time.
4. It compares a baseline with model-based approaches.
5. It uses model refitting to produce a future forecast path.


Statement
1. Show the country-arrivals series together with hotel and stay indicators to position the source-market recovery in a broader tourism context.
2. Use the raw trend, seasonal plot, and decomposition to justify a forecasting approach.
3. Compare Seasonal Naive, ETS, and ARIMA on the same testing window.
4. Conclude with the best model and the forward projection.

## UI Control Mapping

| Parameter | UI Component | Default | Purpose |
|---|---|---|---|
| `series_label` | `selectInput` | Visitor Arrivals: China | choose the target monthly visitor-arrival series by country |
| `horizon` | `sliderInput` | 12 | set the holdout and forward forecast horizon |
| `run_forecast` | `actionButton` | click to run | refresh the forecast after changing controls |

## Output Exposure

| Output | Format | Purpose |
|---|---|---|
| Context comparison chart | `ggplot2` | show whether one country-arrivals series moves in tandem with hotel occupancy, stay length, and room revenue |
| Raw time-series chart | `ggplot2` | inspect the country-level arrival trend and shock/recovery path |
| Seasonal chart | `ggplot2` | compare monthly pattern across years |
| Decomposition plot | `feasts` / `autoplot` | separate trend, seasonality, and remainder |
| Accuracy table | `DT::datatable` | compare testing-set metrics |
| Testing-set forecast plot | `modeltime` plot | assess model behaviour against actual holdout data |
| Forward forecast plot | `modeltime` plot | show the future trajectory after refit |

## Quality Gates

1. The selected series must have at least 24 non-missing monthly observations.
2. The holdout horizon must leave at least 12 points for model fitting.
3. A baseline and at least one model-based forecast must be scored on the same testing window.
4. The forecasting module must keep the shared arrivals workbook as the target backbone while using the tourism workbook only for supporting context indicators.
5. The workflow must include contextual comparison and visual diagnostics before model comparison.
