---
title: "Prototype: Clustering Analysis"
subtitle: "Monthly tourism market states in Singapore's recovery"
format:
  html:
    fig-width: 8
    fig-height: 4.8
execute:
  echo: true
  warning: false
  message: false
---

## 1. Introduction

This prototype follows the same broad storytelling logic as the reference clustering page from Habari Tanzania: introduce the analytical objective, show the code that prepares the data, fit the clustering model, generate the results directly from code, and then explain what those results mean.

The difference is methodological. The Habari Tanzania example works with transformed categorical indicators for latent classes, while this project uses continuous monthly tourism indicators. For that reason, this page uses **standardized k-means clustering** at the **monthly** level.

The analytical question is:

> Can the monthly tourism observations be grouped into a few interpretable market states defined by visitor arrivals, China's market share, hotel occupancy, and length of stay?

All tables and figures below are generated directly from the code chunks in this document.

## 2. Load Packages

### Introduction

The first step is to load the packages needed for Excel import, data transformation, clustering, visualisation, and reporting.

### Code

```{r}
packages <- c("readxl", "dplyr", "ggplot2", "cluster", "knitr")
invisible(lapply(packages, library, character.only = TRUE))

package_table <- data.frame(
  Package = packages,
  Version = vapply(packages, function(pkg) as.character(packageVersion(pkg)), character(1))
)
```

### Result

```{r}
#| echo: false
kable(package_table, align = c("l", "l"))
```

### Explanation

The required packages loaded successfully, so the page can proceed with a fully reproducible workflow.

## 3. Import Data

### Introduction

The clustering module uses the cleaned workbook created for the project. The relevant sheet is `analysis_ready_all`, which already contains the derived variables needed for month-level analysis.

### Code

```{r}
data_candidates <- c(
  file.path("..", "data", "raw", "tourism_four_part_analysis_ready.xlsx"),
  file.path("data", "raw", "tourism_four_part_analysis_ready.xlsx")
)

data_path <- data_candidates[file.exists(data_candidates)][1]
if (is.na(data_path)) {
  stop("Dataset not found. Expected tourism_four_part_analysis_ready.xlsx in data/raw.")
}

tourism_raw <- read_excel(data_path, sheet = "analysis_ready_all") |>
  mutate(date = as.Date(date))

data_import_summary <- data.frame(
  Metric = c(
    "Workbook path",
    "Sheet name",
    "Number of rows",
    "Number of columns",
    "First month",
    "Last month"
  ),
  Value = c(
    data_path,
    "analysis_ready_all",
    nrow(tourism_raw),
    ncol(tourism_raw),
    format(min(tourism_raw$date), "%Y-%m"),
    format(max(tourism_raw$date), "%Y-%m")
  )
)

data_preview <- tourism_raw |>
  select(
    date,
    period,
    visitor_arrivals,
    visitor_arrivals_china,
    china_share,
    hotel_occ,
    avg_stay_monthly_capped
  ) |>
  slice_head(n = 6)
```

### Result

```{r}
#| echo: false
kable(data_import_summary, align = c("l", "l"))
```

```{r}
#| echo: false
kable(data_preview, digits = 3, align = c("l", "l", "r", "r", "r", "r", "r"))
```

### Explanation

The imported dataset contains `r nrow(tourism_raw)` monthly observations from `r format(min(tourism_raw$date), "%Y-%m")` to `r format(max(tourism_raw$date), "%Y-%m")`. The preview confirms that the cleaned sheet already includes the period label, China's market share, and the capped version of monthly stay duration.

## 4. Prepare The Clustering Dataset

### Introduction

The clustering unit must be the month. Following the project plan, the feature set contains:

1. `visitor_arrivals`
2. `china_share`
3. `hotel_occ`
4. `avg_stay_monthly_capped`

The raw `avg_stay_monthly` is kept for interpretation, but the capped version is used for modeling to reduce the influence of extreme pandemic months.

### Code

```{r}
feature_vars <- c(
  "visitor_arrivals",
  "china_share",
  "hotel_occ",
  "avg_stay_monthly_capped"
)

cluster_input <- tourism_raw |>
  select(
    date,
    period,
    visitor_arrivals,
    china_share,
    hotel_occ,
    avg_stay_monthly,
    avg_stay_monthly_capped
  ) |>
  filter(if_all(all_of(feature_vars), ~ !is.na(.x)))

missing_table <- data.frame(
  Variable = feature_vars,
  Missing_values = vapply(feature_vars, function(x) sum(is.na(cluster_input[[x]])), integer(1))
)

period_table <- cluster_input |>
  count(period, name = "Months")

cluster_matrix <- scale(as.matrix(cluster_input[, feature_vars, drop = FALSE]))

standardization_check <- data.frame(
  Variable = colnames(cluster_matrix),
  Mean_after_scaling = round(colMeans(cluster_matrix), 3),
  SD_after_scaling = round(apply(cluster_matrix, 2, sd), 3)
)
```

### Result

```{r}
#| echo: false
kable(missing_table, align = c("l", "r"))
```

```{r}
#| echo: false
kable(period_table, align = c("l", "r"))
```

```{r}
#| echo: false
kable(standardization_check, align = c("l", "r", "r"))
```

### Explanation

The selected modeling fields have no missing values after filtering, leaving `r nrow(cluster_input)` usable months. The period distribution also shows that the dataset spans all three project phases, and the standardization check confirms that each modeling variable now contributes on a comparable scale to the distance calculation.

## 5. Determine The Number Of Clusters

### Introduction

The draft suggested starting with `k = 3`, but that decision still needs evidence. To evaluate the candidate cluster counts, the page computes:

1. the total within-cluster sum of squares for the elbow pattern, and
2. the mean silhouette score for separation quality.

### Code

```{r}
candidate_k <- 2:6
distance_matrix <- dist(cluster_matrix)

model_metrics <- bind_rows(lapply(candidate_k, function(k) {
  set.seed(42)
  fit <- kmeans(cluster_matrix, centers = k, nstart = 25)
  sil <- silhouette(fit$cluster, distance_matrix)

  data.frame(
    k = k,
    tot_withinss = fit$tot.withinss,
    silhouette = mean(sil[, "sil_width"], na.rm = TRUE)
  )
}))

elbow_plot <- ggplot(model_metrics, aes(x = k, y = tot_withinss)) +
  geom_line(color = "#0f3d3e", linewidth = 1.1) +
  geom_point(color = "#d86f45", size = 2.6) +
  scale_x_continuous(breaks = candidate_k) +
  labs(
    x = "Number of clusters (k)",
    y = "Total within-cluster sum of squares"
  ) +
  theme_minimal(base_size = 12)

silhouette_plot <- ggplot(model_metrics, aes(x = factor(k), y = silhouette, fill = k == 3)) +
  geom_col(width = 0.68, show.legend = FALSE) +
  scale_fill_manual(values = c("TRUE" = "#0f6b5b", "FALSE" = "#c7d4cf")) +
  labs(
    x = "Number of clusters (k)",
    y = "Mean silhouette score"
  ) +
  theme_minimal(base_size = 12)

metrics_table <- model_metrics |>
  transmute(
    K = k,
    `Total within-cluster SS` = round(tot_withinss, 1),
    `Mean silhouette` = round(silhouette, 3)
  )
```

### Result

```{r}
#| echo: false
#| label: fig-elbow
#| fig-cap: "Elbow diagnostic for candidate cluster counts."
elbow_plot
```

```{r}
#| echo: false
#| label: fig-silhouette
#| fig-cap: "Mean silhouette score by candidate cluster count."
silhouette_plot
```

```{r}
#| echo: false
kable(metrics_table, align = c("r", "r", "r"))
```

### Explanation

The highest silhouette score occurs at `k = r model_metrics$k[which.max(model_metrics$silhouette)]`, with a mean silhouette of `r round(max(model_metrics$silhouette), 3)`. However, the elbow pattern bends noticeably by `k = 3`, and the three-cluster solution gives a more useful market narrative than a simple two-way split. For that reason, the prototype keeps `k = 3` as the final setting while acknowledging that the choice prioritises interpretability over the absolute best silhouette value.

## 6. Fit The Final K-Means Model

### Introduction

With `k = 3` selected, the next step is to fit the final k-means model, assign each month to a cluster, and then translate the technical clusters into market-state labels.

### Code

```{r}
selected_k <- 3
set.seed(42)
cluster_fit <- kmeans(cluster_matrix, centers = selected_k, nstart = 25)
pca_fit <- prcomp(cluster_matrix)

cluster_results <- cluster_input |>
  mutate(
    cluster_id = cluster_fit$cluster,
    pc1 = pca_fit$x[, 1],
    pc2 = pca_fit$x[, 2]
  )

cluster_profile_raw <- cluster_results |>
  group_by(cluster_id) |>
  summarise(
    months = n(),
    visitor_arrivals = mean(visitor_arrivals),
    china_share = mean(china_share),
    hotel_occ = mean(hotel_occ),
    avg_stay_monthly = mean(avg_stay_monthly),
    avg_stay_monthly_capped = mean(avg_stay_monthly_capped),
    .groups = "drop"
  )

label_map <- cluster_profile_raw |>
  arrange(visitor_arrivals) |>
  mutate(
    state_label = c(
      "Pandemic shock months",
      "Recovery transition months",
      "High-performance months"
    )
  ) |>
  select(cluster_id, state_label)

cluster_results <- cluster_results |>
  left_join(label_map, by = "cluster_id") |>
  mutate(
    state_label = factor(
      state_label,
      levels = c(
        "Pandemic shock months",
        "Recovery transition months",
        "High-performance months"
      )
    ),
    period = factor(period, levels = c("pre_covid", "covid_shock", "recovery"))
  )

cluster_profile <- cluster_profile_raw |>
  left_join(label_map, by = "cluster_id") |>
  mutate(
    state_label = factor(
      state_label,
      levels = c(
        "Pandemic shock months",
        "Recovery transition months",
        "High-performance months"
      )
    )
  ) |>
  arrange(state_label)

cluster_size_table <- cluster_results |>
  count(state_label, name = "Months")

center_table <- as.data.frame(cluster_fit$centers) |>
  mutate(cluster_id = as.integer(rownames(cluster_fit$centers))) |>
  left_join(label_map, by = "cluster_id") |>
  transmute(
    `Market state` = state_label,
    `Center: visitor arrivals` = round(visitor_arrivals, 3),
    `Center: China share` = round(china_share, 3),
    `Center: hotel occupancy` = round(hotel_occ, 3),
    `Center: stay length` = round(avg_stay_monthly_capped, 3)
  ) |>
  arrange(match(`Market state`, levels(cluster_results$state_label)))
```

### Result

```{r}
#| echo: false
kable(cluster_size_table, align = c("l", "r"))
```

```{r}
#| echo: false
kable(center_table, align = c("l", "r", "r", "r", "r"))
```

### Explanation

The final model produces three clearly different groups. The largest group is the **high-performance** state with `r cluster_size_table$Months[cluster_size_table$state_label == "High-performance months"]` months, while the **pandemic shock** and **recovery transition** states capture the disrupted and transitional months that differ strongly from normal market conditions.

The centroid table is shown on the standardized scale, so the signs matter more than the raw magnitudes. Positive values indicate above-average conditions on a feature, while negative values indicate below-average conditions.

## 7. Visualise And Interpret The Results

### Introduction

The next step is to turn the cluster assignments into interpretable outputs: a reduced-dimension scatter plot, a business-friendly profile table, and a timeline showing when each market state appears.

### Code

```{r}
state_palette <- c(
  "Pandemic shock months" = "#b6405d",
  "Recovery transition months" = "#ef9f23",
  "High-performance months" = "#0f6b5b"
)

pca_plot <- ggplot(cluster_results, aes(x = pc1, y = pc2, color = state_label, shape = period)) +
  geom_point(size = 3.1, alpha = 0.88) +
  scale_color_manual(values = state_palette) +
  labs(
    x = "Principal component 1",
    y = "Principal component 2",
    color = "Market state",
    shape = "Period"
  ) +
  theme_minimal(base_size = 12)

timeline_plot <- ggplot(cluster_results, aes(x = date, y = state_label, color = state_label)) +
  geom_point(size = 2.8, alpha = 0.9) +
  scale_color_manual(values = state_palette) +
  labs(
    x = "Month",
    y = "Market state",
    color = "Market state"
  ) +
  theme_minimal(base_size = 12)

profile_table <- cluster_profile |>
  transmute(
    `Market state` = state_label,
    Months = months,
    `Mean visitor arrivals` = format(round(visitor_arrivals), big.mark = ","),
    `Mean China share` = sprintf("%.1f%%", 100 * china_share),
    `Mean hotel occupancy` = sprintf("%.1f%%", hotel_occ),
    `Mean stay (raw)` = round(avg_stay_monthly, 1),
    `Mean stay (capped for model)` = round(avg_stay_monthly_capped, 1)
  )

period_state_view <- as.data.frame.matrix(with(cluster_results, table(state_label, period)))
period_state_view$state_label <- rownames(period_state_view)
rownames(period_state_view) <- NULL
period_state_view <- period_state_view |>
  select(state_label, pre_covid, covid_shock, recovery) |>
  transmute(
    `Market state` = state_label,
    `Pre-COVID` = pre_covid,
    `COVID shock` = covid_shock,
    Recovery = recovery
  )
```

### Result

```{r}
#| echo: false
#| label: fig-pca
#| fig-cap: "Monthly observations projected into two principal components and colored by the inferred market states."
pca_plot
```

```{r}
#| echo: false
kable(profile_table, align = c("l", "r", "r", "r", "r", "r", "r"))
```

```{r}
#| echo: false
#| label: fig-timeline
#| fig-cap: "Timeline of inferred market states across the study window."
timeline_plot
```

```{r}
#| echo: false
kable(period_state_view, align = c("l", "r", "r", "r"))
```

### Explanation

The three-state solution is substantively interpretable.

- **Pandemic shock months** average only `r format(round(cluster_profile$visitor_arrivals[cluster_profile$state_label == "Pandemic shock months"]), big.mark = ",")` arrivals, with mean hotel occupancy around `r sprintf("%.1f%%", cluster_profile$hotel_occ[cluster_profile$state_label == "Pandemic shock months"])`.
- **Recovery transition months** average `r format(round(cluster_profile$visitor_arrivals[cluster_profile$state_label == "Recovery transition months"]), big.mark = ",")` arrivals, showing a partial rebound but not a full return to pre-pandemic market performance.
- **High-performance months** average `r format(round(cluster_profile$visitor_arrivals[cluster_profile$state_label == "High-performance months"]), big.mark = ",")` arrivals and `r sprintf("%.1f%%", cluster_profile$hotel_occ[cluster_profile$state_label == "High-performance months"])` hotel occupancy, making them the closest to normal or strong tourism conditions.

One important nuance is that **China's share can be high even when China's absolute arrival count is still low**. This happens because the share is relative to the whole market. During the most disrupted months, the overall tourism market shrank so sharply that China's proportion could remain elevated despite low absolute volumes. That is why the prototype keeps both volume and share in the interpretation.

The period table also shows the temporal logic of the clusters:

- the **high-performance** state contains all `pre_covid` months and most `recovery` months,
- the **pandemic shock** state is concentrated entirely in the `covid_shock` phase,
- and the **recovery transition** state bridges the end of the shock period and the earliest recovery months.

## 8. UI Design For The Shiny Module

### Introduction

The assignment requires the prototype to define the user interface that will later expose the clustering logic in the Shiny app.

### Code

```{r}
ui_parameter_table <- data.frame(
  Parameter = c("period_filter", "k_value", "random_seed", "run_model", "download_clusters"),
  UI_component = c(
    "checkboxGroupInput",
    "sliderInput",
    "numericInput",
    "actionButton",
    "downloadButton"
  ),
  Default = c("all periods", "3", "42", "click to run", "shown after model run"),
  Purpose = c(
    "filter months by study period",
    "compare alternative cluster counts",
    "keep the result reproducible",
    "avoid unnecessary recomputation",
    "export month-level state assignments"
  )
)

ui_output_table <- data.frame(
  Output = c(
    "state_scatter_plot",
    "state_profile_table",
    "state_timeline_plot",
    "silhouette_value",
    "cluster_assignment_export"
  ),
  Format = c("ggplot2", "DT::datatable", "ggplot2", "value box / text", "CSV download"),
  Purpose = c(
    "show separation between market states",
    "compare cluster averages",
    "link states back to the historical timeline",
    "report model quality",
    "support downstream analysis"
  )
)
```

### Result

```{r}
#| echo: false
kable(ui_parameter_table, align = c("l", "l", "l", "l"))
```

```{r}
#| echo: false
kable(ui_output_table, align = c("l", "l", "l"))
```

### Explanation

This UI design keeps the module focused and interpretable. The controls expose only the parameters that matter for a prototype, while the outputs preserve the analytical story developed in the code above.

## 9. Conclusion

This prototype now follows a stricter reproducible pattern:

1. introduce the analytical objective,
2. show the code,
3. let the code generate the result,
4. and then explain the meaning of that result.

The main conclusion remains the same, but it is now supported more transparently: monthly tourism observations in Singapore do not form one uniform sequence. They separate into interpretable market states associated with visitor volume, China's market share, hotel occupancy, and stay duration.
